---
title: "IDS intestigation worksheet"
author: "by Kate Supremacy: Connor, Kate, Ruth, Torgua & Kate"
date: "`r Sys.Date()`"
output: html_document
---

**Note:** You can use this file as you 'working document' where you can try out various investigation ideas and keep notes about your findings. How you use and structure this file is up to you. It is recommended that you keep notes about what you are investigating and what you find as this will make the process of creating your presentation and report easier. Please note that you _do not_ need to submit this file as part of your group project.



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r load-lib, message = FALSE}
library(tidyverse)
# Add any other libraries here
library(lubridate)
library(tidymodels)

```


```{r load-data}
# load your data 
nyweather <- read_csv("data/ny_weather.csv")

nyweather <- nyweather %>%
  rename(
    "date" = `DATE`,
    "cool_deg_hrs" = `HLY-CLDH-NORMAL`,
    "cloud_brkn_pct" = `HLY-CLOD-PCTBKN`,
    "cloud_clr_pct" = `HLY-CLOD-PCTCLR`,
    "cloud_few_pct" = `HLY-CLOD-PCTFEW`,
    "cloud_ovrcst_pct" = `HLY-CLOD-PCTOVC`,
    "cloud_scat_pct" = `HLY-CLOD-PCTSCT`,
    "dewp_10" = `HLY-DEWP-10PCTL`,
    "dewp_90" = `HLY-DEWP-90PCTL`,
    "dewp_avg" = `HLY-DEWP-NORMAL`,
    "heat_index_avg" = `HLY-HIDX-NORMAL`,
    "heat_deg_hrs" = `HLY-HTDH-NORMAL`,
    "pres_10pctl" = "HLY-PRES-10PCTL",
    "pres_90pctl" = "HLY-PRES-90PCTL",
    "pres_avg" = `HLY-PRES-NORMAL`,
    "temp_10pctl" = "HLY-TEMP-10PCTL",
    "temp_90ptcl" = "HLY-TEMP-90PCTL",
    "temp_avg" = `HLY-TEMP-NORMAL`,
    "wind_chill_avg" = `HLY-WCHL-NORMAL`,
    "wind_main_dir" = "HLY-WIND-1STDIR",
    "wind_main_pct" = "HLY-WIND-1STPCT",
    "wind_2nd_dir" = "HLY-WIND-2NDDIR",
    "wind_2nd_pct" = "HLY-WIND-2NDPCT",
    "wind_spd_avg" = `HLY-WIND-AVGSPD`,
    "pct_calm" = `HLY-WIND-PCTCLM`,
    "wind_vector_dir" = `HLY-WIND-VCTDIR`,
    "wind_vector_spd" = `HLY-WIND-VCTSPD`
  ) %>% # renamed columns
  select(-c("STATION", "NAME")) %>% # got rid of station and name as we know these
  separate(date, into = c("day", "time"), sep = "T") # separated into day and time (might want to change variable types)
#please also note that pressure will not change!! so is not worth looking at

#replacing missing data with nas
nyweather[nyweather == -777.7] <- NA

#correcting percentage columns that are 10 times to big
nyweather<- nyweather %>% 
  mutate(across(c(cloud_brkn_pct,cloud_clr_pct,cloud_few_pct,cloud_ovrcst_pct,cloud_scat_pct, wind_main_pct,wind_2nd_pct), ~ .x/10))

#changing temperature to degrees Celsius because Fahrenheit is completely meaningless to me
nyweather<- nyweather %>% 
  mutate(across(c(dewp_10,dewp_90,dewp_avg,heat_index_avg,temp_10pctl,temp_90ptcl,temp_avg,wind_chill_avg), ~ signif((5/9)*(.x-32), 3)))
```
```{r}
# Which direction is the prevailing wind most frequently in? 
nyweather %>%
  count(wind_main_dir) %>%
  arrange(desc(n)) 
# Using the frequency table and the document in the reference tab of the Google doc, the prevailing wind most frequently in the North West direction.

# Which direction is the secondary wind most frequently in?
nyweather %>%
  count(wind_2nd_dir) %>%
  arrange(desc(n))
# Using the frequency table and the document in the reference tab of the Google doc, the secondary wind most frequently in the West direction.

# Both the prevailing wind and the secondary winds least frequent direction is East.

nyweather %>%
  count(cool_deg_hrs) %>%
  arrange(desc(n)) %>%
  head(5)

#Why are so many of the cooling degree hour entries -777.7? In the document in the reference tab of the Google doc it says, "-7777: a non-zero value that would round to zero". Which I am assuming has a decimal point missing (or added). Do we want to replace these entries with zeros?
```
```{r}
#I am assuming that the percentages of the clouds are multiplied by 10 to get rid of decimals making them "permilleages". I will check they total 1000.
nyweather_test <- nyweather %>%
  mutate(
    total_pct_of_cloud = cloud_brkn_pct + 
      cloud_clr_pct + 
      cloud_few_pct +
      cloud_ovrcst_pct +
      cloud_scat_pct)

nyweather_test %>%
  count(total_pct_of_cloud)

# They don't all total to 1000 which I will assume is though rounding discrepancies.
# Do we want to divide them by ten to make them more recognizable percentages?

# I noticed that two columns look quite similar, I will put them next to each other.
glimpse(select(nyweather, heat_index_avg, temp_avg))

#Also is it in Fahrenheit? Otherwise it is very hot to be in New York! Do we want to convert it to degrees Celsius?
```

```{r}
nyweather %>%
  summarise(
    mean_temp_year = mean(temp_avg),
    median_temp_year = median(temp_avg)
  )
```
```{r}
# Investigating how the average temperature effects the wind direction, more specifically the probability of it being in the west direction (including north west and south west).
nyweather <- nyweather %>%
  mutate(
    wind_main_west = case_when(
      wind_main_dir == "6" ~ "yes",
      wind_main_dir == "7" ~ "yes",
      wind_main_dir == "8" ~ "yes",
      TRUE ~ "no"
    ),
    wind_main_west = fct_relevel(wind_main_west, "no", "yes")
  ) # Creating a column for whether the wind is in the west direction or not.
```

```{r}
set.seed(1234)
nyweather_split <- initial_split(nyweather)
nyweather_train <- training(nyweather_split)
nyweather_test  <- testing(nyweather_split) # Splitting data into training and testing.

wind_dir_rec_1 <- recipe(
  wind_main_west ~ temp_avg,
  data = nyweather_train  # Created a recipe.
)

wind_dir_mod_1 <- logistic_reg() %>%
  set_engine("glm")     # Created a model.

wind_dir_wflow_1 <- workflow() %>%
  add_model(wind_dir_mod_1) %>%
  add_recipe(wind_dir_rec_1) #Added them to a work flow.
```

```{r}
wind_dir_fit_1 <- wind_dir_wflow_1 %>%
  fit(data = nyweather_train)  # Applying work flow to training data.

tidy(wind_dir_fit_1)  # You can see from the table below, that as the average temperature increases the smaller the logic of the probability, and therefore the smaller the probability that the wind will be in the west direction. Or in other words if it is colder it is more likely the wind will be going in the west direction (Including north west and south west direction).

# The formula of the line is log(p/1-p) = 2.63 - 0.17 temp_avg, where p is the probability that the wind is in the west direction.
```

 
```{r}
wind_dir_pred_1 <- predict(wind_dir_fit_1, nyweather_test, type = "prob") %>% 
  bind_cols(nyweather_test) 
wind_dir_pred_1 # Using fitted model to predict the testing data.

wind_dir_pred_1 %>%
  roc_curve(
    truth = wind_main_west,
    .pred_yes,             
    event_level = "second"
  ) %>%
  autoplot() # Plotting a ROC curve

wind_dir_pred_1 %>%
  roc_auc(
    truth = wind_main_west,
    .pred_yes,             
    event_level = "second"
  ) # As seen below the ROC area is relatively close to one which is the ideal value. This means the model has fit the testing data quite well.
```
```{r}
# linear regression of wind chill against heat index
set.seed(1234)
nyweather_split2 <- initial_split(nyweather)
nyweather_train2 <- training(nyweather_split2)
nyweather_test2  <- testing(nyweather_split2)

wind_heat_mod <- linear_reg() %>%
  set_engine("lm")

wind_heat_rec <- recipe(
  wind_chill_avg ~ heat_index_avg,
  data = nyweather_train2
)

wind_heat_wflow <- workflow() %>%
  add_model(wind_heat_mod) %>%
  add_recipe(wind_heat_rec)

wind_heat_fit <- wind_heat_wflow %>%
  fit(data = nyweather_train2)

wind_heat_pred <- predict(wind_heat_fit, nyweather_test2) %>%
  bind_cols(nyweather_test2 %>% select(wind_chill_avg, heat_index_avg))

rsq(wind_heat_pred, truth = wind_chill_avg, estimate = .pred)
rmse(wind_heat_pred, truth = wind_chill_avg, estimate = .pred)
```
```{r}
wind_heat_fit <- wind_heat_wflow %>%
  fit(data = nyweather_train2)
wind_heat_fit_tidy <- tidy(wind_heat_fit) 
wind_heat_fit_aug  <- augment(wind_heat_fit, nyweather_train2) %>%
  mutate(res_cat = ifelse(.resid > 0, TRUE, FALSE))
wind_heat_fit_aug # Created an augmented table with the residuals.

ggplot(data = wind_heat_fit_aug,
       aes(x = heat_index_avg,
           y = wind_chill_avg)) +
  geom_point(aes(x = heat_index_avg, 
                 y = wind_chill_avg,
                 color = res_cat),
             alpha = 0.7) + # Added the points 
  geom_line(aes(x = heat_index_avg, 
                y = .pred), 
            linewidth = 1) + # Added the predication line.
  geom_segment(mapping = aes(x = heat_index_avg, 
                             y = wind_chill_avg,
                             xend = heat_index_avg, 
                             yend = .pred), 
               color = "orange", 
               alpha = 0.4) + # Added line connecting the predicted points to the recorded points.
  labs(
    title = "Wind Chill vs. Heat Index",
    subtitle = "Residual Plot",
    x = "Heat Index (units)",
    y = "Wind Chill (units)"
  ) + # Added labels
  guides(color = FALSE) + # Remove key
  scale_color_manual(values = c("red", "forestgreen")) # Whether residuals are positive are negative.
```
```{r}
ggplot(data = wind_heat_fit_aug,
       aes(x = heat_index_avg,
           y = wind_chill_avg)) +
  geom_point(aes(x = heat_index_avg, 
                 y = wind_chill_avg,
                 color = res_cat),
             alpha = 0.7) +
  geom_line(aes(x = heat_index_avg, y = .pred), linewidth = 0.75) +
  geom_point(mapping = aes(x = heat_index_avg,
                           y = .pred), 
             color = "orange",
             alpha = 0.7)+
  geom_segment(mapping = aes(x = heat_index_avg, 
                             y = wind_chill_avg,
                             xend = heat_index_avg, 
                             yend = .pred), 
               color = "orange", 
               alpha = 0.4) + 
  labs(
    title = "Wind Chill vs. Heat Index",
    subtitle = "Residual Plot",
    x = "Heat Index (units)",
    y = "Wind Chill (units)"
  ) +
  guides(color = FALSE) +
  scale_color_manual(values = c("maroon", "forestgreen")) + 
  coord_cartesian(
  xlim = c(3, 8),
  ylim = c(0, 5)) # Same graph again just zoomed into specific point.
```
```{r}
ggplot(wind_heat_fit_aug, mapping = aes(x = .pred, y = .resid)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, 
             color = "gray", 
             lty = "dashed") +
  labs(x = "Predicted Wind Chill", 
       y = "Residuals") # Residual Plot, unfortunately there looks like there is another relationship going on.
```


